Future challenges for Research Software Engineering

### 2025-03-27 future_challenges_for_RSEng.tex
Add section "Impact of Complexity and Usability in Research Software"
Co-authored-by: Jan Linxweiler <jlinx@users.noreply.github.com>

% Article frontmatter
% ===================
\title{Future challenges for Research Software Engineering} % Title of the article
%\short{} % Short title of the article (optional)
\author{
Florian Goth\texorpdfstring{\autref{1}}{},
Leyla Jael Castro\texorpdfstring{\autref{1}}{},
Gerasimos Chourdakis\texorpdfstring{\autref{1}}{},
Simon Christ\texorpdfstring{\autref{1}}{},
Jeremy Cohen\texorpdfstring{\autref{1}}{},
Jean-Noël Grad\texorpdfstring{\autref{1}}{},
Magnus Hagdorn\texorpdfstring{\autref{1}}{},
Toby Hodges\texorpdfstring{\autref{1}}{},
Jan Linxweiler\texorpdfstring{\autref{1}}{},
Frank Löffler\texorpdfstring{\autref{1}}{},
Michele Martone\texorpdfstring{\autref{1}}{},
Jan Philipp Thiele\texorpdfstring{\autref{1}}{},
Nick del Grosso\texorpdfstring{\autref{1}}{}
} % Authors and references to addresses
\institute{Fantasy University\autlabel{1}} % Institutes with labels
\abstract{Admittedly, it’s still only 2025 and we are actually stillworking on establishing Research Software Engineering
as a new domain with the hope of ensuring that it gets the recognition it deserves.
And while it is certainly too early to already herald the end of RSEng,
we should acknowledge that the future is in constant flux and
it is worthwhile to discuss how RSEng will/could change in response to these upcoming challenges in an open discussion.} % Abstract of the article
\keywords{Keywords go here.} % Keywords for the article

@@

\subsection{Where has abstraction brought us, and which complexities are now hidden behind it?}
FG-NOTE: What would be great, if we can put it in a black box in our daily work and not worry about it any longer?
FG-NOTE: Here we would need a catchy word, that I think should exist out there. Is it Interface? Encapsulation?
Is there a word for "abstracting sth. away?"

\subsection{Impact of Increased Usability in Research Software}

>>>>

% Article frontmatter
% ===================
\title{Future challenges for Research Software Engineering} % Title of the article
%\short{} % Short title of the article (optional)
\author{
Florian Goth\texorpdfstring{\autref{1}}{},
Leyla Jael Castro\texorpdfstring{\autref{1}}{},
Gerasimos Chourdakis\texorpdfstring{\autref{1}}{},
Simon Christ\texorpdfstring{\autref{1}}{},
Jeremy Cohen\texorpdfstring{\autref{1}}{},
Jean-Noël Grad\texorpdfstring{\autref{1}}{},
Magnus Hagdorn\texorpdfstring{\autref{1}}{},
Toby Hodges\texorpdfstring{\autref{1}}{},
Jan Linxweiler\texorpdfstring{\autref{1}}{},
Frank Löffler\texorpdfstring{\autref{1}}{},
Michele Martone\texorpdfstring{\autref{1}}{},
Jan Philipp Thiele\texorpdfstring{\autref{1}}{},
Nick del Grosso\texorpdfstring{\autref{1}}{}
} % Authors and references to addresses
\institute{\autlabel{1} Fantasy University} % Institutes with labels
\abstract{Admittedly, it’s still only 2025 and we are actually stillworking on establishing Research Software Engineering
as a new domain with the hope of ensuring that it gets the recognition it deserves.
And while it is certainly too early to already herald the end of RSEng,
we should acknowledge that the future is in constant flux and
it is worthwhile to discuss how RSEng will/could change in response to these upcoming challenges in an open discussion.} % Abstract of the article
\keywords{Keywords go here.} % Keywords for the article

@@

\subsection{Impact of Complexity and Usability in Research Software}

The workshop participants found that in earlier learning stages, details may be intentionally hidden, fearing a diminishing capability to handle technical complexity in research environments. As a result, RSEs invest more effort in making individual research software components less complex with better user experience. This made participants worry about growing mismatch of peoples' expectations. Participants identified the over-reliance on generative AI tools as an extreme manifestation of this issue. No concrete examples were given on which complexities are now hidden behind abstractions.

We acknowledge but would also like to challenge the observations of the participants. We believe that ``we are standing on the shoulders of giants'': By reducing the complexity of individual software components, we are now able to develop systems of complexity and capability not previously possible, allowing the scientific community to address questions not previously in grasp. When it comes to teaching, it is true that students need to learn how to handle complexity, but it is also important to learn how to abstract from details (create black boxes) and build hierarchical systems from those abstractions and thereby manage complexity.
